Reduced growth due to belowground sink limitation is not fully explained by reduced photosynthesis

### 2016-06-09 manuscript/manuscript_pve.Rmd
possibly still an issue with Nmass calcultaion
changed labels for clairty and made other thesis reviewer comments
ongoing revisions from BM

---
title: Reduced growth due to belowground sink limitation is not explained by reduced photosynthesis
author: ""
date: ""
output:
  word_document:
    reference_docx: manuscriptstyle_newphyt.docx
  pdf_document: default
  html_document:
    number_sections: yes
csl: new-phytologist.csl
bibliography: references2.bib
---

@@

- The difficulty in quantifying all components of plant carbon (C) balance inhibits our ability to interpret growth limitation in woody species. This study manipulated soil volume to test how growth is coupled to physiology, C allocation, and sink activity in *Eucalyptus tereticornis* seedlings.

- Seedlings were grown in a range of container sizes and containers were planted flush to the soil alongside freely-rooted (free) seedlings. We developed a growth model that utilized net leaf photosynthesis rates (*A*~n~) to allocate daily C uptake towards mass growth of stems, leaves and roots.

- Container volume negatively affected aboveground growth throughout the experiment, yet, mass partitioning to leaves, stems and roots was unchanged after 120 days. Photosynthetic capacity was significantly reduced in containered seedlings, and was related to reduced leaf nitrogen content and starch accumulation. Although belowground sink limitation reduced *A*~n~, the model concluded that these reductions were too small to explain observed growth responses.

- As *A*~n~ and growth were not tightly coordinated, an unaccounted-for pool of photosynthetic C resulted in seedlings with soil volume restriction. These results highlight the need to utilize mass balance approaches when evaluating plant C allocation and confirm that important feedbacks exist between belowground sink strength and leaf C uptake.

@@

#Introduction
Understanding plant growth and its relationship to carbon (C) assimilation requires knowledge of the mass balance that must be achieved between C uptake and subsequent allocation to growth, storage, and respiration. As woody plants have highly integrated systems of competing carbohydrate sinks [@kozlowski1992carbohydrate], growth is driven by the allocation of photosynthate from source leaves among these different tissue sinks. At long enough time scales leaf photosynthesis (*A*~n~) and respiratory losses together determine net C balance and plant growth. At shorter temporal scales growth can instead be mediated by tissue C storage pools [@chapin1990ecology], which disrupts the coordination between *A*~n~ and growth. This discrepancy has led to the current debate on how strongly plant growth is controlled by either source or sink activity. As plant growth cannot be simply determined by *A*~n~, growth models must take into account several simultaneous processes affecting the C balance at any given time [@fourcaud2008plant]. Whether growth limitation in woody species is driven mostly by *A*~n~ or other components of the C balance remains largely uncertain.

\
One approach to examine the coordination of *A*~n~ and growth is to manipulate C source activity. Examples include elevated CO~2~ and partial defoliation experiments. Elevated CO~2~ has been shown to increase *A*~n~ [@drake1997more; @ainsworth2007response]; across four forest Free-Air CO~2~ Enrichment (FACE) experiments this increased photosynthetic uptake resulted in a stimulation of biomass production by 23 % on average [@norby2005forest]. Evidence from a wide range of elevated CO~2~ experiments, however, also reveals that even with an average photosynthetic enhancement of over 30 %, the biomass growth rate only increases by around 10 % [@kirschbaum2011does]. In partial defoliation experiments, increases in *A*~n~ of the remaining foliage are commonly shown, and are attributed to reduction in end-product inhibition [@iglesias2002regulation; @zhou2003changes; @handa2005test], enhanced biochemical activity [@ovaska1993b;@layne1995end;@pinkard2011gas], increased stomatal conductance [@layne1995end], enhanced leaf nutrient status [@turnbull2007increased;@pinkard2011gas] and regulatory sugar signaling [@eyles2013whole]. Similar to elevated CO~2~ experiments, these increases in *A*~n~ in defoliation experiments did not always produce increased growth due to reductions in meristem sink strength [@palacio2012fast], C limitation to mycorrhizal colonization [@markkola2004defoliation], or an overall decrease in whole plant C gain [@ovaska1993a]. These manipulations of C source activity demonstrate that changes in *A*~n~ do not always correspond with proportional responses of growth.

\
Alternatively, manipulation of plant tissue C sinks is often used to investigate the correlation between *A*~n~ and growth. Plant metabolic signalling networks have the ability to regulate photosythetic activity, according to carbon status of different plant tissues [@paul2001sink]. If sink inhibition of *A*~n~ occurs, a close coordination between declines in *A*~n~ and growth should be expected. The occurrence of photosynthetic down regulation has been tested through fruit removal and phloem girdling. In these studies, down regulation of *A*~n~ was found to correlate with carbohydrate accumulation resulting from reduced tissue sink strength  [@iglesias2002regulation;@urban2007girdling; @haouari2013fruit]. However, reductions in *A*~n~ were also attributed to biochemical limitations prior to carbohydrate accumulation [@nebauer2011photosynthesis], irreversible photo-oxidative damage [@duan2008photosynthetic] and stomatal limitation [@li2005photosynthesis]. This lack of continuity among findings is not surprising, as we still know little about the balance between assimilation, storage and growth across temporal scales in plants [@smith2007coordination]. As these manipulations likely impact source as well as sink activity simultaneously, affect water transport, are very extreme, or are specific to the occurrence of large fruiting sinks, they tell us little about source-sink coordination in typical growing conditions for woody species.

\
An alternative experimental approach is to reduce belowground C sink strength in tree seedlings by manipulating rooting volume, by varying the container size [@arp1991effects; @nesmith1998effect; @poorter2012pot]. Inadequate rooting volume can decrease C sink strength by progressively restricting root growth in growing plants [@thomas1991root]. Container size studies frequently exhibit photosynthetic down-regulation, likely as a result of sink limitation [@arp1991effects; @mcconnaughay1991physical; @gunderson1994photosynthetic; @sage1994acclimation; @maina2002intra; @ronchi2006growth]. A meta-analysis by Poorter et al. [-@poorter2012pot] concluded that *A*~n~ is the process likely to be the strongest affected by pot size and may best explain the effects on biomass seen in the large number of studies where containers are used. This conclusion arises because plants grown in small containers are shown to accumulate leaf starch while having lower assimilation rates [@robbins1988effect]. However, evidence in support for a trade-off between C storage and growth in trees is, to date, inconclusive [@palacio2014does].  Whether down regulation of *A*~n~ and subsequent reductions of growth are quantitavely similiar has rarely been tested. Based on these previous studies, using a range of container sizes as a sink-strength manipulation provdies a way to empirically test the extent to which growth and *A*~n~ are coordinated.

\
This study utilizes a novel field design to investigate the coordination between growth and *A*~n~ in *Eucalyptus tereticornis* Sm. seedlings, by manipulating container size and thus rooting volume. Seedlings were maintained under well watered conditions in order to isolate the effect of restricted soil volume. We used freely-rooted seedlings as a control for the container size treatments. We tracked leaf-level gas exchange, leaf carbohydrate accumulation and seedling allometry over the course of 4 months. Empirical results were combined with a simple plant growth model to simulate seedling growth with a C mass balance approach, which was then compared to observed harvested seedling mass. The model used whole-plant C gain, scaled from instantaneous rates of leaf *A*~n~, to quantify seedling dry mass production over the 120 day experiment.

\
Manipulations of container size were expected to induce a belowground sink limitation compared to free seedlings, resulting in growth reductions in the smallest containers. Specifically, reductions in biomass partitioning to root mass, relative to tree size, were predicted to occur with decreasing rooting volume and soil nutrients in container treatments. We also expected a correlation between carbohydrate accumulation and photosynthetic capacity as a function of soil volume, as reduced sink strength would lead to accumulation of leaf non-structural carbohydrates and down regulation of *A~n~*. Taking into account measured rates of photosynthesis, we expected biomass predictions from the growth model to correspond to observed seedling mass.

@@

\
Leaf total non-structural carbohydrate (TNC) concentration was analyzed on dried and milled leaf samples using a total starch assay kit (Megazyme International, Wicklow, Ireland) and includes the starch  and soluble sugar concentrations (mg g^-1^). Starch was quantified using a thermostable $\alpha$-amylase and amyloglucosidase assay [@McCleary_starch] and soluble sugars were determined following the anthrone method [@ebell1969variation]. Complete methods of the TNC assay are described in @mitchell2013drought. TNC-free specific leaf area (SLA~f~, m^2^ kg^-1^) for leaves sampled during gas exchange campaigns, was calculated by first subtracting the TNC content from individual dry leaf mass before dividing leaf area by leaf mass. Similarly, TNC-free leaf N content (N~f~, %) was calculated on all gas exchange leaves from leaf mass without TNC content.

##Seedling growth model
We developed a simple seedling growth model that utilized leaf *A*~n~ rates to allocate daily C assimilate towards biomass production of stems, leaves, fine roots and coarse roots. The model begins with mean initial tissue component biomass (leaf~0~, stem~0~ and root~0~) and a starting leaf area (L~0~) measured prior to planting. The initial biomass of roots was divided evenly between fine and coarse roots. The daily net biomass production of seedlings (*G*), from @makela1997carbon, is then given by

@@

##Growth and morphology
Plant height, diameter and leaf area diverged between container volumes soon after start of the experiment (Figure `r allomfig`a-c). First to diverge was seedling leaf area, which significantly differed among soil volumes (*P* < 0.029) during the 5^th^ week of the experiment. Significant differences were then observed in both height (8^th^ week) and then diameter (9^th^ week) significantly among soil volumes (*P* < 0.045 & 0.035, respectively). The large reductions in height gain and total leaf area in smaller compared to larger containers continued throughout the experiment. In this field study, colder temperatures and reductions in total PPFD per day (Figure `r airvarsfig`) likely slowed the growth of seedlings in the final weeks of the experiment. Seedlings maintained diameter growth throughout the experiment, although growth was marginal in smaller soil volumes in the final month. Final seedling height significantly increased with increasing soil volume (*P* < 0.001). Increases in both final stem diameter and cumulative leaf area (both *P* < 0.001) were found with increasing soil volume and these differences were driven mainly by the largest container and the free seedling treatments.

@@

\
Neither $\Psi$~pd~ nor $\Psi$~l~ were different across treatments, with mean values of -0.27 and -1.2 MPa across all seedlings, respectively. Although *g*~s~ in free seedlings was generally higher than those in containers (Table `r pvetab3`, *P* < 0.001), the mean rates for all seedlings were high at 0.37 mol m^-2^ s^-1^ and did not change throughout the course of the experiment. Additionally, leaf $\delta$^13^C at final harvest was not different across treatments (Table `r pvetab1`). Combined, these indices provide strong evidence that water stress was not apparent on these seedlings throughout the experiment. Soil N at harvest was not different across soil volumes ($\bar{x}$ = 4.5 %), with minimal decreases from pre-planting value ($\bar{x}$ = 4.9 %). This similarity indicates that nutrient leaching from free seedlings or from draining of containers following natural rainfall events did not differ between treatments.

@@

#Discussion
This study utilized a novel field design to manipulate belowground sink limitation and physically restrict *Eucalyptus tereticornis* seedling biomass production. We addressed questions regarding the coordination of *A*~n~ and growth by complementing empirical results with a C mass balance model. We found that reductions in leaf *A*~n~ across container sizes, when integrated across the 120 day experiment, were alone insufficient to account for observed reductions in total plant biomass production.

##Reductions in growth and physiology under sink limitation
Soon after seedlings became established, both height and diameter growth were negatively affected by decreasing soil volume. Low growth led to large reductions in biomass in small containers, compared to freely rooted seedlings. We analyzed the relationship between biomass growth and soil volume and found an increase of 34 % with a doubling of container volume, consistent with the meta-analysis of Poorter et al. [-@poorter2012pot]. These growth reductions were expected, as the impedance of root growth can cause reductions in overall plant growth and activity [@mcconnaughay1991physical; @young1997mechanical]. It has been shown that roots subjected to environmental stress may send inhibitory signals to the shoots that affect *g*~s~, cell expansion, cell division and the rate of leaf appearance [@passioura2002soil]. Here, we observed a large divergence in leaf area between seedlings in containers and free seedlings through time, with the eventual cessation of new leaf growth in seedlings in small containers.

\
Decreases in *A*~sat~ occurred simultaneously with reductions growth of seedlings in containers. This observation initially suggests that sink limitation may be driving the down regulation of photosynthesis, and thus growth. However, there are several possible mechanisms that can explain reduced *A*~n~ in small containers including reduced water availability or nutrient content, or reduced sink strength [@poorter2012pot]. It was therefore necessary to examine each of these factors to determine if the induced belowground sink limitation actually triggered photosynthetic down regulation.

@@

\
With respect to nutrient availability, reductions in *A*~n~ were positively correlated with decreases in leaf N~f~ and leaf N~f~ was considerably reduced for seedlings in containers. Since reductions in leaf N~f~ were detected with TNC-free leaf mass, TNC dilution of leaf N was accounted for in all seedlings. It it possible that either physical root restriction or decreased supply reduced seedling N uptake in small containers. Although root N was on average higher in free seedlings at the end of the experiment, it was not consistently higher than the smallest container volume treatment. Unrestricted mycorrhizal recruitment could have instead facilitated the increases in leaf N~f~ in free seedlings, but this effect is unknown. Soil N also declined evenly across all treatments, providing no clear evidence for decreased supply between free seedlings and seedlings in containers. In these already low fertility soils, it is possible that seedlings in containers simply grew into increasing N limitation which negatively affected belowground sink strength. Although no clear feedback could be determined between the available soil N pool and decreases in leaf N~f~, the effects of belowground sink limitation on *A*~n~ of seedlings in containers was evident throughout the experiment.

@@

##Biomass partitioning under sink limitation
As biomass partitioning is likely controlled by the source and sink strength of all organs [@poorter2012biomass], it was important to determine which tissue components were most affected by the container size treatments. It was necessary to distinguish if growth was affected beyond ontogenetic constraints, by correcting for size, as biomass distribution is strongly size-dependent [@gould1966allometry; @lleonart2000removing]. In this study, there was no significant difference in root, leaf, or stem biomass partitioning with reduced soil volume compared to free seedlings, outside of ontogenetic drift (Figure `r allocfig`a,b). This result is unexpected as shifts in partitioning have been noted specifically for nutrient limitation [@mcconnaughay1999biomass, and references therein]. Surprisingly, a constant ratio of fine root mass to leaf mass was observed across all treatments suggesting a functional partitioning response to optimize resource gain did not occur.

@@

\
Focusing only on empirical results in this study points to *A*~n~ driving reductions in seedling growth. Our modelling results reveal that other components of the C balance, beyond *A*~n~, are required to explain the observed seedling biomass response to sink manipulation. This discrepancy is noteworty, as partial accounting of the different mechaninsms involved in plant carbon balance can lead to erroneous conclusions [@valentini2000respiration]. For example, the fraction of photosynthate used in respiration is known to vary depending on species and local environmental conditions [@Lambers2008], yet is often considered a static parameter in process based growth models. The use of fixed rates of respiration in our model likely underestimates seedling responses to sink limitation. Thus, our modelling results agree with conclusions from Valentine and M&#228;kel&#228; [-@valentine2005bridging] where the problem with predicting tree growth is a problem in forecasting the assimilation and allocation of C and other constituents. Future empirical and modelling studies should focus on how feedbacks from sink activity affect both rates of *A*~n~ and the fate of C allocated to growth, respiration and C storage in leaves. However, the degree to which these mechanisms regulate the C available for growth will undoubtedly shift across different species and environmental conditions.

>>>>

---
title: Reduced growth due to belowground sink limitation is not fully explained by reduced photosynthesis
author: ""
date: ""
output:
  word_document:
    reference_docx: manuscriptstyle_newphyt.docx
  pdf_document: default
  html_document:
    number_sections: yes
csl: new-phytologist.csl
bibliography: references2.bib
---

@@

- Sink limitation is known to reduce plant growth, but it is not known how plant carbon (C) balance is affected, limiting our ability to predict growth under sink-limited conditions.

- We manipulated soil volume to impose sink limitation of growth in Eucalyptus tereticornis seedlings. Seedlings were grown in containers of different sizes and planted flush to the soil alongside freely-rooted seedlings.

- Container volume negatively affected aboveground growth throughout the experiment, but mass partitioning to leaves, stems and roots was unchanged after 120 days. Photosynthetic capacity was significantly reduced in containerized seedlings, and was related to both reduced leaf nitrogen content and starch accumulation. However, the reduction in net leaf photosynthesis over the growth period was insufficient to explain the reduction in growth, so that we also observed a reduction in whole-plant carbon use efficiency (CUE) with container size.

- Our results show that sink limitation affects plant growth through feedbacks to both photosynthesis and CUE. Mass balance approaches to predicting plant growth under sink-limited conditions need to incorporate both of these feedbacks.

@@

#Introduction
The source-sink limitation paradigm for plant growth argues that growth may be limited either by source activity, i.e. the amount of carbon that the plants are able to take up through photosynthesis, or by sink strength, i.e. the amount of carbon that the plants are able to utilize in growth.  To be able to predict plant growth responses to environmental change, we need models that can account for both source and sink limitation.

Most current models of plant growth are based on the physical principle of mass balance: carbon (C) entering the plant through photosynthesis must be balanced by C loss to respiration and C allocation to growth and storage. Plant growth is thus predicted as the outcome of the processes of photosynthesis, respiration and allocation.  It is easy to see how source-limited growth can be implemented in such models: increasing source activity increases photosynthetic C uptake, flowing through to increased growth. It is less easy to know how to implement sink limitation of growth in such models, because it requires that we know whether, and how, sink limitation influences the contributing processes of photosynthesis, respiration and allocation. It has been suggested that sink limitation affects the carbon balance by causing a down-regulation of photosynthetic rate, possibly mediated by carbohydrate accumulation (refs), but there is also evidence that photosynthesis and growth are decoupled under sink limitation (refs), suggesting that other processes must also be affected.

In the absence of a good understanding of the effects of sink limitation on physiological processes, most attempts to model sink limitation of growth to date have done so by abandoning the principle of mass balance (e.g. Leuzinger; CLM nutrient limitation; Fatichi). This approach is unsatisfactory because mechanistic models should obey the laws of physics. Our goal in this paper was to investigate how the carbon balance of plants is modified by sink limitation, by experimentally imposing a sink limitation to growth.

Understanding plant growth and its relationship to carbon (C) assimilation requires knowledge of the mass balance that must be achieved between C uptake and subsequent allocation to growth, storage, and respiration. As woody plants have highly integrated systems of competing carbohydrate sinks [@kozlowski1992carbohydrate], growth is driven by the allocation of photosynthate from source leaves among these different tissue sinks. At long enough time scales leaf photosynthesis (*A*~n~) and respiratory losses together determine net C balance and plant growth. At shorter temporal scales growth can instead be mediated by tissue C storage pools [@chapin1990ecology], which disrupts the coordination between *A*~n~ and growth. This discrepancy has led to the current debate on how strongly plant growth is controlled by either source or sink activity. As plant growth cannot be simply determined by *A*~n~, growth models must take into account several simultaneous processes affecting the C balance at any given time [@fourcaud2008plant]. Whether growth limitation in woody species is driven mostly by *A*~n~ or other components of the C balance remains largely uncertain.

\
In previous studies, the sink limitation of growth has been explored by manipulations of both C source and C sink activity. Examples of experimental approaches that manipulate C source activity include elevated atmospheric CO2 concentration and partial defoliation experiments.

Elevated CO~2~ causes an increase in leaf-level photosynthetic rate, *A*~n~ [@drake1997more; @ainsworth2007response]; across four forest Free-Air CO~2~ Enrichment (FACE) experiments this increased photosynthetic uptake resulted in a stimulation of biomass production by 23 % on average [@norby2005forest]. Evidence from a wide range of elevated CO~2~ experiments, however, also reveals that even with an average photosynthetic enhancement of over 30 %, the biomass growth rate only increases by around 10 % [@kirschbaum2011does].

In partial defoliation experiments, source activity is decreased by reduction of photosynthesizing leaf area. In these experiments, compensatory increases in *A*~n~ of the remaining foliage are commonly found, and have been attributed to several processes including indirect effects on resource availability, such as increaed stomatal conductance [@layne1995end] or enhanced leaf nutrient status [@turnbull2007increased; @pinkard2011gas], or more direct effects of modified sink-source balance such as, reduction in end-product inhibition [@iglesias2002regulation; @zhou2003changes; @handa2005test], enhanced biochemical activity [@ovaska1993b;@layne1995end;@pinkard2011gas], and regulatory sugar signaling [@eyles2013whole]. However, there have been few attempts *(any??)* to determine the overall C balance in these experiments, and it is unclear whether the change in growth following defoliation can be explained by the reduction in leaf area and changes to A~n~

Similar to elevated CO~2~ experiments, these increases in *A*~n~ in defoliation experiments did not always produce increased growth due to reductions in meristem sink strength [@palacio2012fast], C limitation to mycorrhizal colonization [@markkola2004defoliation], or an overall decrease in whole plant C gain [@ovaska1993a]. These manipulations of C source activity demonstrate that changes in *A*~n~ do not always correspond with proportional responses of growth.

\
Alternatively, sink limitation can be investigated by manipulating plant tissue C sinks, for example by fruit remove or phloem girdling. In these studies, down regulation of *A*~n~ is often observed, and typically correlates with carbohydrate accumulation resulting from reduced tissue sink strength  [@iglesias2002regulation;@urban2007girdling; @haouari2013fruit]. However, reductions in *A*~n~ have also been attributed to biochemical limitations prior to carbohydrate accumulation [@nebauer2011photosynthesis], irreversible photo-oxidative damage [@duan2008photosynthetic] and stomatal limitation [@li2005photosynthesis]. This lack of continuity among findings is not surprising, as we still know little about the balance between assimilation, storage and growth across temporal scales in plants [@smith2007coordination]. As these manipulations likely impact source as well as sink activity simultaneously, affect water transport, are very extreme, or are specific to the occurrence of large fruiting sinks, they tell us little about source-sink coordination in typical growing conditions for woody species.

\
Yet another experimental approach is to reduce belowground C sink strength in tree seedlings by manipulating container size and thus rooting volume [@arp1991effects; @nesmith1998effect; @poorter2012pot]. Inadequate rooting volume can decrease C sink strength by progressively restricting root growth in growing plants [@thomas1991root].

Container size studies frequently exhibit photosynthetic down-regulation, likely as a result of sink limitation [@arp1991effects; @mcconnaughay1991physical; @gunderson1994photosynthetic; @sage1994acclimation; @maina2002intra; @ronchi2006growth]. A meta-analysis by Poorter et al. [-@poorter2012pot] concluded that *A*~n~ is the process likely to be the strongest affected by pot size and may best explain the effects on biomass seen in the large number of studies where containers are used. This conclusion arises because plants grown in small containers are shown to accumulate leaf starch while having lower assimilation rates [@robbins1988effect]. However, evidence in support for a trade-off between C storage and growth in trees is, to date, inconclusive [@palacio2014does].  Whether down regulation of *A*~n~ and subsequent reductions of growth are quantitavely similiar has rarely been tested. Based on these previous studies, using a range of container sizes as a sink-strength manipulation provdies a way to empirically test the extent to which growth and *A*~n~ are coordinated.

\
Our goal in this paper was to impose a sink limitation of growth, by manipulating container size and thus rooting volume, and to examine how the C balance of the plants was affected. We grew *Eucalyptus tereticornis* seedlings in a range of container sizes, using freely-rooted seedlings as a control for the container size treatments. Seedlings were maintained under well-watered conditions in order to isolate the effect of restricted soil volume from that of low water availability. We tracked leaf-level gas exchange, leaf carbohydrate accumulation and seedling allometry over the course of 4 months. We drove a simple whole-plant photosynthesis model with measured leaf gas exchange and interpolated leaf area data to estimated net canopy carbon uptake over the experimental period, and compared the result with total seedling biomass. We anticipated that both photosynthesis and growth would be reduced under belowground sink limitation. The principal questions that we addressed were: Firstly, which physiological process best explains the down-regulation of photosynthesis? Secondly, does this down-regulation fully explain the observed response of growth, or are other components of the carbon balance also affected

@@

\
Leaf total non-structural carbohydrate (TNC) concentration was analyzed on dried and milled leaf samples using a total starch assay kit (Megazyme International, Wicklow, Ireland) and includes the starch  and soluble sugar concentrations (mg g^-1^). Starch was quantified using a thermostable $\alpha$-amylase and amyloglucosidase assay [@McCleary_starch] and soluble sugars were determined following the anthrone method [@ebell1969variation]. Complete methods of the TNC assay are described in @mitchell2013drought. TNC-free specific leaf area (SLA~f~, m^2^ kg^-1^) for leaves sampled during gas exchange campaigns, was calculated by first subtracting the TNC content from individual dry leaf mass before dividing leaf area by leaf mass. Similarly, TNC-free leaf N concentration (N~f~, %) was calculated on all gas exchange leaves from leaf mass without TNC content.

##Net carbon uptake model
We developed a simple seedling growth model that utilized leaf *A*~n~ rates to allocate daily C assimilate towards biomass production of stems, leaves, fine roots and coarse roots. The model begins with mean initial tissue component biomass (leaf~0~, stem~0~ and root~0~) and a starting leaf area (L~0~) measured prior to planting. The initial biomass of roots was divided evenly between fine and coarse roots. The daily net biomass production of seedlings (*G*), from @makela1997carbon, is then given by

@@

##Growth and morphology
Plant height, diameter and leaf area diverged between container volumes soon after start of the experiment (Figure `r allomfig`a-c). First to diverge was seedling leaf area, which significantly differed among soil volumes (*P* < 0.029) during the 5^th^ week of the experiment. Significant differences were then observed in both height (8^th^ week) and then diameter (9^th^ week) significantly among soil volumes (*P* < 0.045 & 0.035, respectively). The large reductions in height gain and total leaf area in smaller compared to larger containers continued throughout the experiment. In this field study, colder temperatures and reductions in total PPFD per day (Figure `r airvarsfig`) likely slowed the growth of seedlings in the final weeks of the experiment. Seedlings maintained diameter growth throughout the experiment, although growth was marginal in smaller soil volumes in the final month. Final seedling height significantly increased with increasing soil volume (*P* < 0.001). Increases in both final stem diameter and cumulative leaf area (both *P* < 0.001) were found with increasing soil volume. We did not find a continuous effect of container size; rathe, plant size was similar among smaller containers, and these differences were driven mainly by the largest container and the free seedling treatments.

@@

\
Neither $\Psi$~pd~ nor $\Psi$~l~ were different across treatments, with mean values of -0.27 and -1.2 MPa across all seedlings, respectively. Although *g*~s~ in free seedlings was generally higher than those in containers (Table `r pvetab3`, *P* < 0.001), the mean rates for all seedlings were high at 0.37 mol m^-2^ s^-1^ and did not change throughout the course of the experiment. Additionally, leaf $\delta$^13^C at final harvest was not different across treatments (Table `r pvetab1`). Combined, these indices provide strong evidence that water stress was not apparent on these seedlings throughout the experiment. Soil N at harvest was not different across soil volumes ($\bar{x}$ = .045 %), with minimal decreases from pre-planting values ($\bar{x}$ = .049 %). This similarity indicates that nutrient leaching from free seedlings or from draining of containers following natural rainfall events did not differ between treatments.

@@

#Discussion
This study utilized a novel field design to manipulate belowground sink limitation and physically restrict *Eucalyptus tereticornis* seedling biomass production. Our goal was to identify how sink limitation of growth modified the carbon balance of seedlings. We found a strong reduction of growth in containerized plants. Leaf *A*~n~ was also strongly reduced, but this reduction was insufficient to fully explain the reduction in growth, implying that whole-plant respiratory loss was likely affected as well.

We addressed questions regarding the coordination of *A*~n~ and growth by complementing empirical results with a C mass balance model. We found that reductions in leaf *A*~n~ across container sizes, when integrated across the 120 day experiment, were alone insufficient to account for observed reductions in total plant biomass production.

##Changes in growth and physiology under sink limitation
Soon after seedlings became established, both height and diameter growth were negatively affected by decreasing soil volume. Low growth rates led to large reductions in biomass in small containers, compared to freely rooted seedlings. We analyzed the relationship between biomass growth and soil volume and found an increase of 34 % with a doubling of container volume, consistent with the meta-analysis of Poorter et al. [-@poorter2012pot]. These growth reductions were expected, as the impedance of root growth can cause reductions in overall plant growth and activity [@mcconnaughay1991physical; @young1997mechanical]. It has been shown that roots subjected to environmental stress may send inhibitory signals to the shoots that affect *g*~s~, cell expansion, cell division and the rate of leaf appearance [@passioura2002soil]. Here, we observed a large divergence in leaf area between seedlings in containers and free seedlings through time, with the eventual cessation of new leaf growth in seedlings in small containers.

\
Decreases in *A*~sat~ occurred simultaneously with reductions of growth of seedlings in containers. This observation initially suggests that sink limitation may be driving the down regulation of photosynthesis, and thus growth. However, there are several possible mechanisms that can explain reduced *A*~n~ in small containers including reduced water availability or nutrient content, or reduced sink strength [@poorter2012pot]. It was therefore necessary to examine each of these factors to determine the mechanism by which the induced belowground sink limitation triggered photosynthetic down regulation.

@@

\
With respect to nutrient availability, reductions in *A*~n~ were positively correlated with decreases in leaf N~f~ and leaf N~f~ was considerably reduced for seedlings in containers. Since reductions in leaf N~f~ were detected with TNC-free leaf mass, TNC dilution of leaf N was accounted for in all seedlings. It is possible that either physical root restriction or decreased supply reduced seedling N uptake in small containers. Although root N was on average higher in free seedlings at the end of the experiment, it was not consistently higher than the smallest container volume treatment. Unrestricted mycorrhizal recruitment could have instead facilitated the increases in leaf N~f~ in free seedlings, but this effect is unknown. Soil N also declined evenly across all treatments, providing no clear evidence for decreased supply between free seedlings and seedlings in containers. In these already low fertility soils, it is possible that seedlings in containers simply grew into increasing N limitation which negatively affected belowground sink strength. Although no clear feedback could be determined between the available soil N pool and decreases in leaf N~f~, the effects of belowground sink limitation on *A*~n~ of seedlings in containers was evident throughout the experiment.

@@

##Biomass partitioning under sink limitation
As biomass partitioning is likely controlled by the source and sink strength of all organs [@poorter2012biomass], it was important to determine which tissue components were most affected by the container size treatments. It was necessary to distinguish if growth was affected beyond ontogenetic constraints, by correcting for size, as biomass distribution is strongly size-dependent [@gould1966allometry; @lleonart2000removing]. In this study, there was no significant difference in root, leaf, or stem biomass partitioning with reduced soil volume compared to free seedlings, once ontogenetic drift was taken into account (Figure `r allocfig`a,b). This result is unexpected as shifts in partitioning have been noted specifically for nutrient limitation [@mcconnaughay1999biomass, and references therein]. Surprisingly, a constant ratio of fine root mass to leaf mass was observed across all treatments suggesting a functional partitioning response to optimize resource gain did not occur.

@@

\
Focusing only on empirical results in this study points to *A*~n~ driving reductions in seedling growth. Our modelling results reveal that other components of the C balance, beyond *A*~n~, are required to explain the observed seedling biomass response to sink manipulation. This discrepancy is noteworthy, as partial accounting of the different mechanisms involved in plant carbon balance can lead to erroneous conclusions [@valentini2000respiration]. For example, the fraction of photosynthate used in respiration is known to vary depending on species and local environmental conditions [@Lambers2008], yet is often considered a static parameter in process based growth models. The use of fixed rates of respiration in our model likely underestimates seedling responses to sink limitation. Thus, our modelling results agree with conclusions from Valentine and M&#228;kel&#228; [-@valentine2005bridging] where the problem with predicting tree growth is a problem in forecasting the assimilation and allocation of C and other constituents. Future empirical and modelling studies should focus on how feedbacks from sink activity affect both rates of *A*~n~ and the fate of C allocated to growth, respiration and C storage in leaves. However, the degree to which these mechanisms regulate the C available for growth will undoubtedly shift across different species and environmental conditions.
